---
title: "Hw 2"
author: "Amanda Kramer"
date: 9/30/20
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(knitr)
```

## Problem 1

First I will read and clean the Mr. Trash Wheel sheet:

*   specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
*   use reasonable variable names
*   omit rows that do not include dumpster-specific data
*   round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

```{r}
trash_wheel = 
  read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  select(-x15, -x16, -x17) %>% 
  mutate(sports_balls = as.integer(sports_balls))
```

Next I will read and clean precipitation data for 2017 and 2018. For each omit rows without precipitation data and add a variable year. Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

```{r}
precip_17 = 
  read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 6, skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = 2017) %>% 
  select(year, month, total)

precip_18 = 
  read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5, skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = 2018) %>% 
  select(year, month, total)

precip_all = 
  full_join(precip_17, precip_18) %>% 
  mutate(month = month.name[month])
```

```{r include = FALSE}
tw_sb =
  trash_wheel %>%
  filter(year == 2017) %>%
  select(year, sports_balls)
```

My trash_wheel data set has `r nrow(trash_wheel)` observations, and my precip_all data set has `r nrow(precip_all)` observations. Trash_wheel includes variables to show the number of items removed from the harbor, such as `glass_bottles` and `grocery_bags`. We can also see that the median number of `sports_balls` in the dumpster in 2017 was `r median(pull(filter(trash_wheel, year==2017), sports_balls))`. The precip_all dataframe can show us that the total precipitation in 2018 was `r sum(pull(precip_18, total))` inches.